What Makes Us Happy? Examining the Impact of Social, Economic, and Personal Factors

### 2024-12-02 paper/paper.qmd
make changes

---
title: "What Makes Us Happy? Unveiling the Impact of Social, Economic, and Personal Factors"
author: "Shuheng (Jack) Zhou"
thanks: "The GitHub Repository containing all data, R code, and other files used in this project is located here:<https://github.com/Shuhengzhou03/Factors-Influencing-Happiness.git>"
date: today
date-format: long
abstract: "This study employs a logistic regression model to analyze the impact of various social, economic, and personal factors on happiness. The research focuses on key variables such as income, marital status, education level, job satisfaction, and the number of children, examining their influence on individuals' self-reported happiness levels. The results highlight the relative importance of these factors, providing insights into the complex relationship between socioeconomic conditions and personal well-being while identifying the most significant predictors of happiness."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

Happiness is a fundamental aspect of human well-being, yet understanding the complex factors that contribute to an individual's happiness has long remained a challenge in both psychological and sociological research. While much attention has been given to the impact of socio-economic factors such as income, education, and employment on happiness, less focus has been directed towards understanding the role of personal characteristics and socio-cultural variables in shaping happiness. This paper seeks to fill this gap by examining how different variables, such as marital status, job satisfaction, income, and education, influence an individual’s happiness.

The core of this research lies in analyzing a dataset that includes a wide range of personal and socio-economic factors. These factors include marital status, number of children, job satisfaction, income, and education level, along with personal characteristics such as age and gender. By applying a Bayesian logistic regression model, this study quantifies the independent impact of each predictor variable on happiness, providing a deeper understanding of which factors have the most significant influence on happiness.

Our findings reveal that job satisfaction and marital status play an important role in determining happiness, with individuals who report higher satisfaction in these areas more likely to report being "very happy." Additionally, income and education level also emerge as significant contributors, with higher income and education levels positively influencing happiness, though the impact varies across different social groups.

The significance of these findings extends beyond academic interest, offering practical insights for policymakers and social scientists interested in enhancing well-being. This research contributes to a deeper understanding of how different socio-economic factors influence individual happiness, providing valuable guidance for social policy development.

This paper is structured to provide a thorough examination of the variables influencing happiness. Following the introduction in @sec-intro, @sec-data outlines the dataset used in the study, detailing the sources of the data and the selection of key variables for analysis. @sec-model introduces the Bayesian logistic regression model employed to analyze the relationships between the predictors and happiness, providing a statistical framework for understanding how various factors independently contribute to happiness. @sec-result presents the findings from the Bayesian model, explaining how job satisfaction, marital status, and other factors impact the likelihood of being "very happy." @sec-discussion explores the implications of these findings, discusses the limitations of the study, and suggests avenues for future research to better understand the factors influencing happiness. @sec-model-details provides detailed plots of posterior predictive checks and model diagnostics, ensuring the robustness of our findings.

@@

The estimand of this study is the probability of an individual self-reporting as "very happy." Since happiness is a subjective experience influenced by a wide range of complex factors, it is practically challenging to conduct a comprehensive survey of the happiness levels of the entire population. Therefore, this study utilizes sample data from the **GSS Data Explorer** and applies a Bayesian logistic regression model to estimate this probability. The sample data includes variables such as marital status, job satisfaction, income, education level, age, gender, and the number of children, representing both personal characteristics and socio-economic factors.

Through this model, the study aims to quantify the independent impact of each variable on happiness, assessing their contributions to individual happiness while controlling for other factors. The findings not only provide deeper insights into the determinants of happiness but also offer empirical support for policies aimed at enhancing societal well-being.

@@

The data used in this study was sourced from the [GSS Data Explorer](https://gssdataexplorer.norc.org/MyGSS) [@gssdataexplorer], an online platform maintained by NORC that provides a wide range of resources from the General Social Survey (GSS). The GSS dataset includes extensive information on social, economic, and personal characteristics, making it an essential tool for studying social behaviors and trends. This study utilized relevant data downloaded from the platform and performed filtering and preprocessing to focus on factors influencing happiness.

@@

- **`tidyverse`** [@tidyverse]: Provided a comprehensive set of tools for data manipulation and visualization, significantly simplifying the workflow.
- **`palmerpenguins`** [@palmerpenguins]: Offered example datasets and tools, aiding in the quick testing of analysis code.
- **`broom`** [@broom]: Used for tidying model outputs, making them easier to integrate and interpret.
- **`ggplot2`** [@ggplot2]: Provided powerful and flexible data visualization capabilities for creating charts tailored to the study's requirements.
- **`dplyr`** [@dplyr]: Facilitated efficient data manipulation and transformation, serving as a core tool for data cleaning and preparation.
- **`tidyr`** [@tidyr]: Used to reshape and organize data, enabling effective analysis and visualization.
- **`arrow`** [@arrow]: Efficiently read and wrote large datasets, enhancing data processing performance.
- **`scales`** [@scales]: Improved chart readability by formatting scales and labels to enhance visual presentation.
- **`rstanarm`** [@rstanarm]: Simplified Bayesian modeling, providing an intuitive interface for complex Bayesian analysis.
- **`brms`** [@brms]: A flexible modeling tool built on Stan, used to perform comprehensive Bayesian regression analysis on the data.
- **`bayesplot`**[@bayesplot]: A visualization package for Bayesian model diagnostics, posterior checks, and MCMC outputs, offering tools for clear and intuitive graphical summaries.

@@

@tbl-happiness_preview presents the first six rows from the cleaned dataset, focusing on socio-economic and personal variables that influence happiness. The dataset includes key information such as marital status, job satisfaction, income, education level, age, gender, and the number of children, providing a comprehensive basis for analyzing the factors that contribute to individual happiness.

@@

**childs:** A numerical variable indicating the number of children an individual has (e.g., 2). "This variable records the total number of children reported by a respondent. It provides insight into family size and is used to analyze its potential influence on various aspects of well-being, including happiness levels."

@@

**degree:** A categorical variable indicating an individual's highest level of educational attainment (e.g., "Bachelor's"). "This variable classifies respondents into different education categories, such as 'Less than High School,' 'High School,' 'Bachelor's,' 'associate/junior college,' and 'Graduate.' It provides insight into the respondent's educational background and is used to analyze its potential impact on various aspects of life, including happiness levels."

@@

```
@fig-Degree illustrates the relationship between education level and happiness levels, with data points grouped by educational categories for clarity. The bars represent the proportion of happiness levels across different education levels, with distinct shades of green used to differentiate happiness categories. The chart highlights that individuals with higher education levels, such as graduate degrees, tend to report slightly higher levels of happiness, while other education levels show more balanced distributions. The use of proportional stacking and color differentiation aids in visually comparing trends across education levels.

@@

**satjob:** A categorical variable indicating an individual's level of job satisfaction (e.g., "Very Satisfied"). "This variable classifies respondents into different job satisfaction categories, such as 'Very Satisfied,' 'Moderately Satisfied,' 'A Little Dissatisfied,' and 'Very Dissatisfied.' It provides insight into the respondent's feelings about their job and is used to analyze its potential impact on various aspects of well-being, including happiness levels."

@@

```
@fig-Satjob illustrates the relationship between job satisfaction and happiness levels, with data points grouped by satisfaction categories for clarity. The bars represent the proportion of happiness levels across different job satisfaction levels, with distinct shades of green used to differentiate happiness categories. The chart highlights that individuals who are more satisfied with their jobs tend to report higher levels of happiness, while those with lower satisfaction show a more diverse distribution. The use of proportional stacking and color differentiation helps in visually comparing the trends across job satisfaction levels.

@@

**realrinc:** A numerical variable indicating an individual's real income (e.g., "50000"). "This variable measures the respondent's actual income adjusted for inflation, providing a more accurate representation of purchasing power. It is used to classify respondents into different income levels and analyze its potential impact on various aspects of well-being, including happiness levels. Real income offers valuable insights into the economic conditions of respondents and their correlation with life satisfaction."

@@

The predictors in the model include marital status, number of children, age, education level, gender, job satisfaction, and real income. The Bayesian model was fitted using the `brm` function from the brms package [@brms] in R. The model uses a Bernoulli family with a logit link, and priors were specified as $\text{Normal}(0, 2)$ for coefficients and $\text{Cauchy}(0, 2)$ for the intercept, reflecting weakly informative prior beliefs about the effects of each predictor.

@@

The Bayesian Logistic Regression model was chosen for its ability to incorporate prior knowledge and quantify uncertainty in predictions, allowing for a nuanced understanding of the relationships between predictors and happiness. By using weakly informative priors, the model remains flexible while ensuring stable estimation of coefficients. This approach is particularly valuable for exploring individual-level happiness data, where complex interactions and varying effects are expected across predictors.

## Model Summary
We summarized the results of the Bayesian Logistic Regression model using the `summary` function in R, which provides detailed information about the estimated coefficients, their associated uncertainty, and credible intervals. The coefficients indicate the direction and magnitude of the relationship between predictors and the likelihood of being "very happy."

Additionally, we performed posterior predictive checks using the `pp_check` function to evaluate the fit of the model. These checks demonstrated a reasonable agreement between the predicted and observed values, suggesting that the model effectively captures the patterns in the data.

The estimated coefficients revealed significant associations between happiness levels and predictors such as job satisfaction, marital status, and real income. For example, higher job satisfaction and income were associated with an increased likelihood of being "very happy." The credible intervals provided insights into the uncertainty of these estimates, ensuring a robust interpretation.

We also calculated the average predicted probability of being "very happy" across the dataset, which highlights the overall effectiveness of the predictors in explaining variations in happiness levels. These results provide meaningful insights into the factors that contribute to individual happiness.

@@

As detailed in **@tbl-happiness_model_summary**, the coefficient summary provides quantitative insights into the socio-economic and demographic factors influencing individual happiness. For example, the estimated coefficient for **Marital Status: Married** is notably positive (Mean = 0.80), indicating that individuals who are married have a higher likelihood of reporting greater happiness compared to those in the reference category (e.g., never married).

Conversely, the coefficient for **Number of Children** is slightly negative (Mean = -0.10), suggesting that having more children is associated with a marginal decrease in happiness, potentially reflecting the increased responsibilities and financial pressures that come with larger families. Similarly, the negative coefficient for **Age** (Mean = -0.02) suggests a modest decline in happiness with age.

Positive coefficients for **Education Level: Graduate** (Mean = 0.50) and **Job Satisfaction: Moderately Satisfied** (Mean = 0.70) highlight the strong association between higher education, job satisfaction, and increased happiness. Additionally, **Real Income** (Mean = 0.02) shows a small but positive effect, indicating that higher income is correlated with greater happiness, albeit at a modest rate.

Finally, the Intercept (Mean = -1.50) establishes the baseline log-odds for happiness, setting the reference point for interpreting the influence of these predictors. Together, these coefficients illuminate the multifaceted factors shaping happiness, underscoring the importance of socio-economic stability and personal well-being.

@@

# Plot the 90% credible intervals for all model coefficients
# This visualizes the impact of each predictor on the ordered priority levels
mcmc_intervals(posterior, prob = 0.9) +
  labs(
    title = "Happiness Model Coefficients",
    subtitle = "90% credible intervals for the predictors",
    x = "Coefficient Estimate",
    y = "Parameters"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5), # Center the main title
    plot.subtitle = element_text(hjust = 0.5) # Center the subtitle
  )
```
As shown in @fig-happiness_model_coefficients, the Bayesian logistic regression model visualizes the impact of predictor variables on happiness. Each point represents the posterior mean estimate, and the horizontal lines indicate the 90% credible intervals. Key findings include:

- The variable **`Marital Status: Married`** shows a positive coefficient, indicating that married individuals report significantly higher happiness compared to the reference group (e.g., never married).
- **`Education Level: Graduate`** and **`Job Satisfaction: Moderately Satisfied`** exhibit strong positive effects, highlighting the importance of higher education and job satisfaction in boosting happiness.
- Conversely, **`Number of Children`** and **`Age`** show negative coefficients, suggesting that an increase in the number of children and aging slightly reduce happiness levels.
- **`Real Income`** displays a modest positive effect, indicating that higher income is associated with increased happiness, although the effect is relatively small.

These findings provide detailed insights into the socio-economic and demographic factors influencing individual happiness.

@@

This paper explores the factors influencing individual happiness using a Bayesian logistic regression model to analyze self-reported happiness levels. By leveraging a carefully selected subset of socio-economic and demographic variables from a comprehensive dataset, this study identifies patterns that highlight key determinants of happiness and their relative importance in shaping well-being outcomes.

@@

This study carefully selects key variables—marital status, job satisfaction, education level, real income, age, and number of children—based on their strong association with individual happiness. The selection process was guided by both theoretical significance and empirical evidence, ensuring the model accurately reflects real-world influences on well-being. The quantitative analysis demonstrates how these variables play a pivotal role in shaping personal happiness. For instance, marital status and job satisfaction emerge as strong predictors, with married individuals and those who are highly satisfied with their jobs reporting significantly higher levels of happiness.

Interestingly, the analysis reveals a nuanced relationship between the number of children and happiness. While having children can bring joy, larger family sizes are often associated with slightly lower levels of happiness, possibly due to increased financial and caregiving responsibilities. Similarly, higher levels of education and income are positively associated with happiness, illustrating the role of socio-economic stability in improving life satisfaction. These findings not only align with established theories but also provide a deeper understanding of the complex interactions among social, economic, and demographic factors that drive individual well-being.

@@

By selecting variables with strong associations to individual happiness—such as marital status, job satisfaction, education level, income, age, and number of children—this study underscores the multifaceted nature of well-being. The inclusion of these variables reflects both theoretical relevance and empirical robustness, providing insights into the dynamics of happiness. For example, focusing on the positive impact of marriage and job satisfaction aligns with broader sociological findings about emotional and economic stability.

Interestingly, while other variables, such as detailed occupational categories or regional distinctions, were excluded to streamline the analysis, their potential impact opens avenues for future research. Future investigations could explore how specific professions or geographic contexts influence happiness, shedding light on how cultural and environmental factors interact with socio-economic predictors. This targeted approach not only refines current models but also lays the groundwork for exploring the broader complexities of what drives individual happiness.

@@

While this study provides valuable insights into the factors influencing individual happiness, its scope is limited by the variables included and the dataset size. The exclusion of variables such as region of residence or specific occupational categories, though aimed at simplifying the model, restricts the depth of analysis. Future research could incorporate these variables to better understand the interplay of geographical, cultural, and professional contexts on happiness.

Moreover, the model could be enhanced by exploring additional dimensions, such as health conditions or leisure time, to provide a more holistic view of well-being. Integrating data from broader samples across different regions or including longitudinal datasets could improve the model’s generalizability and offer insights into how happiness evolves over time.

Another avenue for future exploration involves examining the impact of external socio-economic factors, such as economic downturns or public policy changes, on happiness. For example, understanding how financial crises or access to social services influence happiness levels could inform policies aimed at improving societal well-being.

Additionally, the use of advanced machine learning techniques could help uncover hidden patterns and interactions among variables that traditional regression models might miss. For instance, exploring non-linear relationships between age, income, and happiness could yield deeper insights into the life stages most conducive to well-being.

These research directions not only build upon the findings of this study but also provide opportunities to address its limitations, contributing to a more comprehensive understanding of the determinants of happiness in diverse contexts.

@@

The potential for future research based on this dataset is vast. One promising avenue involves investigating the intersection between happiness and broader societal factors, such as access to healthcare or the availability of social support systems. By integrating external datasets, future studies could explore how these contextual elements amplify or mitigate the effects of individual-level predictors like income or job satisfaction.

Another intriguing possibility is examining how happiness evolves across different life stages, using longitudinal data to track changes over time. Such research could illuminate critical periods where interventions might have the greatest impact on well-being.

Additionally, incorporating geographical data could provide a deeper understanding of regional variations in happiness, allowing for an analysis of how cultural or environmental factors shape well-being. The dataset’s richness opens the door to these and many other avenues for advancing our understanding of the complex drivers of happiness.

@@

This research highlights the analytical value of understanding the factors that drive individual happiness, such as the significant role of marital status and job satisfaction. These findings can inform policy development and workplace strategies aimed at improving well-being. The positive correlation between higher education levels and happiness also emphasizes the broader societal value of investing in accessible and equitable education systems, reflecting the importance of socio-economic stability in fostering life satisfaction.

>>>>

---
title: "What Makes Us Happy? Examining the Impact of Social, Economic, and Personal Factors"
subtitle: "Marital Status and Job Satisfaction as Major Positive Predictors of Happiness: A Bayesian Analysis of Socio-Economic and Personal Factors Influencing Well-Being"
author: "Shuheng (Jack) Zhou"
thanks: "The GitHub Repository containing all data, R code, and other files used in this project is located here:<https://github.com/Shuhengzhou03/Factors-Influencing-Happiness.git>"
date: today
date-format: long
abstract: "This study employs a logistic regression model to investigate the influence of diverse social, economic, and personal factors on self-reported happiness. By focusing on variables such as income, marital status, education level, job satisfaction, and the number of children, it examines their respective roles in shaping individual well-being. The findings underscore the significance of these factors, illustrating how socioeconomic conditions and personal attributes interact to influence happiness. This analysis offers a clearer perspective on the primary predictors of well-being, with meaningful implications for enhancing both individual and societal happiness."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

Happiness is a cornerstone of human well-being and a long-standing focus of psychological and sociological inquiry. Despite extensive research, understanding the complex interplay of factors contributing to happiness remains a significant challenge. While prior studies have primarily centered on socio-economic determinants such as income, education, and employment, comparatively less attention has been devoted to the influence of personal characteristics and socio-cultural variables. This paper seeks to address this issue by examining how a combination of socio-economic and personal factors collectively shape individual happiness.

The essential role of socio-economic stability in fostering happiness is well-documented. For example, Layard [@article] emphasizes income and job security as fundamental determinants of life satisfaction, while Easterlin [@easterlin1974] demonstrated that rising income does not always translate into increased happiness—a phenomenon widely recognized as the "Easterlin Paradox." More recently, Helliwell et al. [@helliwell2020] have highlighted the importance of social connections and subjective well-being, suggesting that these dimensions often outweigh purely economic metrics. Despite these valuable contributions, a notable issue persists in understanding how socio-economic factors interact with personal characteristics—such as marital status, gender, and the number of children—in shaping happiness.

This study advances the existing literature by integrating socio-economic and demographic variables into a unified analytical framework to quantify their independent and combined effects on happiness. Specifically, the research investigates key variables, including marital status, job satisfaction, income, education, gender, age, and the number of children, utilizing a Bayesian logistic regression model. By focusing on the interplay between personal and socio-economic factors, the study provides a more detailed perspective, addressing important gaps in the literature.

The findings indicate that job satisfaction and marital status are the strongest predictors of happiness, with individuals reporting high satisfaction in these areas significantly more likely to describe themselves as "very happy." Additionally, income and education levels positively correlate with happiness, though their impact varies across demographic groups. Notably, the relationship between the number of children and happiness is complex; while larger families can bring joy, they often introduce financial and caregiving challenges.

These results validate earlier findings on the importance of socio-economic stability while offering new insights into how demographic factors influence well-being. The study provides practical recommendations for policymakers and social scientists to design targeted interventions aimed at enhancing happiness across different population segments.

This paper is structured as follows: @sec-intro introduces the context and objectives of the study. @sec-data outlines the dataset and describes the selection of key variables. @sec-model presents the Bayesian logistic regression framework employed in the analysis. @sec-result discusses the findings, focusing on the effects of job satisfaction, marital status, and other predictors. @sec-discussion explores the broader implications, limitations, and potential directions for future research, while @sec-model-details provides detailed diagnostics and validation to ensure the robustness of the results.

@@

The estimand of this study is the probability of an individual self-reporting as "very happy." Since happiness is a subjective experience influenced by a wide range of complex factors, it is practically challenging to conduct a detailed survey of the happiness levels of the entire population. Therefore, this study utilizes sample data from the GSS Data Explorer and applies a Bayesian logistic regression model to estimate this probability. The sample data includes variables such as marital status, job satisfaction, income, education level, age, gender, and the number of children, representing both personal characteristics and socio-economic factors.

Through this model, the study aims to quantify the independent impact of each variable on happiness, assessing their contributions to individual happiness while controlling for other factors. The findings not only provide a better understanding of the determinants of happiness but also offer empirical support for policies aimed at enhancing societal well-being.

@@

The data utilized in this study was sourced from the [GSS Data Explorer](https://gssdataexplorer.norc.org/MyGSS) [@gssdataexplorer], an online platform maintained by NORC that provides an extensive repository of information collected through the General Social Survey (GSS). The GSS dataset captures a wide range of social, economic, and personal characteristics, making it a useful resource for analyzing societal behaviors and trends. For this study, the primary focus was on variables directly related to individual happiness, such as marital status, job satisfaction, income, education, gender, age, and the number of children.

In addition to the GSS dataset, there are other datasets that could have been used to analyze the factors influencing happiness. For example, the European Social Survey (ESS) provides rich data on life satisfaction and happiness, but it primarily focuses on European countries, making it less suitable for this study's focus on American society. In contrast, the GSS dataset is centered on the U.S. population, offering more relevant data for examining the impact of socio-economic conditions and personal characteristics on happiness in the United States. Similarly, the World Values Survey (WVS) offers global social-economic data, but due to its broad scope, it may not be as suitable for focusing on happiness factors specific to the U.S. population. The American Time Use Survey (ATUS) primarily focuses on time allocation rather than directly measuring life satisfaction or happiness, making it less aligned with the objectives of this study. The GSS dataset was chosen for its focus on U.S. residents and its detailed records of socio-economic and personal characteristics, which are crucial for analyzing the factors influencing happiness among Americans.

The process of transforming real-world phenomena into data entries begins with the GSS survey design, where participants are asked a series of standardized questions designed to capture their demographic and socio-economic characteristics, as well as subjective well-being. For example:

- Happiness was measured using responses to the question, *"Taking all things together, would you say you are very happy, pretty happy, or not too happy?"* Responses were coded as categorical variables in the dataset.

- Income data reflects respondents' self-reported annual income, adjusted for inflation and recorded as a continuous variable.

- Marital status, job satisfaction, education, and other personal attributes were captured through categorical responses, subsequently encoded into numerical or factor variables for analysis.

To ensure the data's relevance to this study, preprocessing steps were undertaken. These included filtering the dataset to include only respondents who had complete entries for all variables of interest and standardizing continuous variables such as age and income to improve interpretability in the modeling process. For instance, missing data for happiness or income was excluded, and categorical variables were re-coded into binary or ordinal formats where applicable.

@@

- **`tidyverse`** [@tidyverse]: Provided a suite of tools for data manipulation and visualization, significantly simplifying the workflow.
- **`palmerpenguins`** [@palmerpenguins]: Offered example datasets and tools, aiding in the quick testing of analysis code.
- **`broom`** [@broom]: Used for tidying model outputs, making them easier to integrate and interpret.
- **`ggplot2`** [@ggplot2]: Provided powerful and flexible data visualization capabilities for creating charts tailored to the study's requirements.
- **`dplyr`** [@dplyr]: Facilitated efficient data manipulation and transformation, serving as a core tool for data cleaning and preparation.
- **`tidyr`** [@tidyr]: Used to reshape and organize data, enabling effective analysis and visualization.
- **`arrow`** [@arrow]: Efficiently read and wrote large datasets, enhancing data processing performance.
- **`scales`** [@scales]: Improved chart readability by formatting scales and labels to enhance visual presentation.
- **`rstanarm`** [@rstanarm]: Simplified Bayesian modeling, providing an intuitive interface for complex Bayesian analysis.
- **`brms`** [@brms]: A flexible modeling tool built on Stan, used to perform Bayesian regression analysis on the data.
- **`bayesplot`**[@bayesplot]: A visualization package for Bayesian model diagnostics, posterior checks, and MCMC outputs, offering tools for clear and intuitive graphical summaries.

@@

@tbl-happiness_preview presents the first six rows from the cleaned dataset, focusing on socio-economic and personal variables that influence happiness. The dataset includes key information such as marital status, job satisfaction, income, education level, age, gender, and the number of children, providing a detailed basis for analyzing the factors that contribute to individual happiness.

@@

**childs:** A numerical variable indicating the number of children an individual has (e.g., 2). "This variable records the total number of children reported by a respondent. It shows family size and is used to analyze its potential influence on various aspects of well-being, including happiness levels."

@@

**degree:** A categorical variable indicating an individual's highest level of educational attainment (e.g., "Bachelor's"). "This variable classifies respondents into different education categories, such as 'Less than High School,' 'High School,' 'Bachelor's,' 'associate/junior college,' and 'Graduate.' It examines the respondent's educational background and is used to analyze its potential impact on various aspects of life, including happiness levels."

@@

```
@fig-Degree illustrates the relationship between education level and happiness levels, with data points grouped by educational categories for clarity. The bars display the proportion of happiness levels by education categories, using distinct shades of green to represent happiness levels. The chart indicates that individuals with higher education, such as graduate degrees, are more likely to report higher happiness levels, while other education categories exhibit more uniform distributions. Proportional stacking and color differentiation help compare trends effectively across education levels.

@@

**satjob:** A categorical variable indicating an individual's level of job satisfaction (e.g., "Very Satisfied"). "This variable categorizes respondents by their job satisfaction levels: 'Very Satisfied,' 'Moderately Satisfied,' 'A Little Dissatisfied,' and 'Very Dissatisfied.' It indicates respondents' attitudes toward their jobs and is analyzed for its influence on happiness levels."

@@

```
@fig-Satjob illustrates the relationship between job satisfaction and happiness levels, with data points grouped by satisfaction categories for clarity. The bars display the proportion of happiness levels by job satisfaction categories, using distinct shades of green to represent happiness levels. The chart indicates that individuals with higher job satisfaction are more likely to report greater happiness, while those with lower satisfaction exhibit a broader distribution. Proportional stacking and color differentiation facilitate comparisons across job satisfaction levels.

@@

**realrinc:** A numerical variable indicating an individual's real income (e.g., "50000"). "This variable represents respondents' inflation-adjusted income, accurately reflecting their purchasing power. It classifies respondents into income levels and analyzes its impact on happiness. Real income captures economic conditions and their association with life satisfaction."

@@

In constructing the model, the predictors selected are directly aligned with the variables discussed in the data section. For example, marital status is included as it is a key socio-economic factor influencing happiness, with prior research showing its strong relationship to well-being. Number of children is included as it can introduce both positive and negative effects on happiness, depending on familial dynamics and financial responsibilities. This variable was treated as a continuous predictor, capturing the direct impact of family size, rather than categorizing it into bins, to retain the detailed effects of the number of children.

Age was treated as a continuous variable, rather than categorizing it into specific age groups, in order to retain more detailed information and reflect the continuous nature of age's impact on happiness, as supported by prior literature. This decision avoids the loss of variability that might occur with grouped age ranges. Similarly, education level and job satisfaction are included as categorical predictors to capture distinct group differences while preserving important socio-economic distinctions that are likely to influence happiness.

The inclusion of gender was based on its significant impact on well-being outcomes in previous research, allowing us to analyze how gender dynamics interact with other socio-economic factors. Real income is included as a continuous variable, as it reflects an individual's economic stability and directly correlates with happiness. The decision to treat income as continuous rather than categorizing it into income brackets helps retain detailed information about the effects of varying income levels on happiness, particularly since the impact of income on well-being may be non-linear.

We employed a Bayesian Logistic Regression model using the `brm` function from the `brms` package [@brms] in R, which is particularly useful for fitting complex regression models with a probabilistic framework. The model assumes a Bernoulli family with a logit link, suitable for binary outcome variables like happiness. Weakly informative priors were applied to the coefficients (`Normal(0, 2)`) and intercept (`Cauchy(0, 2)`), reflecting our prior belief that the effects of each predictor are not strongly deviating from zero, providing flexibility for the model to learn from the data.

In summary, the predictors were selected based on their relevance to understanding individual happiness, with each variable being included to reflect real-world influences as thoroughly as possible. The model’s specification, particularly the treatment of continuous vs. categorical variables, ensures that the relationships between predictors and happiness are captured with as much detail as the data allow.

@@

The Bayesian Logistic Regression model was chosen for its ability to incorporate prior knowledge and quantify uncertainty in predictions, allowing for a detailed understanding of the relationships between predictors and happiness. By using weakly informative priors, the model remains flexible while ensuring stable estimation of coefficients. This approach is particularly valuable for exploring individual-level happiness data, where complex interactions and varying effects are expected across predictors.

## Model Summary

We summarized the results of the Bayesian Logistic Regression model using the summary function in R, which provides detailed information about the estimated coefficients, their associated uncertainty, and credible intervals. The coefficients indicate the direction and magnitude of the relationship between predictors and the likelihood of being "very happy."

Additionally, we performed posterior predictive checks using the pp_check function to evaluate the fit of the model. These checks indicated a reasonable agreement between the predicted and observed values, suggesting that the model effectively captures the patterns in the data.

The estimated coefficients revealed significant associations between happiness levels and predictors such as job satisfaction, marital status, and real income. For example, higher job satisfaction and income were associated with an increased likelihood of being "very happy." The credible intervals provided findings about the uncertainty of these estimates, ensuring a robust interpretation.

We also calculated the average predicted probability of being "very happy" across the dataset, which highlights the overall effectiveness of the predictors in explaining variations in happiness levels. These results provide meaningful findings about the factors that contribute to individual happiness.

@@

## Coefficient Summary {#sec-coefficient-summary}

As detailed in @tbl-happiness_model_summary, the coefficient summary offers a quantitative perspective on the socio-economic and demographic determinants of individual happiness. For instance, the estimated coefficient for **Marital Status: Married** is significantly positive (Mean = 0.80), indicating that married individuals are more likely to report higher levels of happiness compared to those in the reference category (e.g., never married). This highlights the role of marital status in fostering emotional and social well-being.

In contrast, the coefficient for **Number of Children** is slightly negative (Mean = -0.10), suggesting that larger family sizes are associated with a marginal decrease in happiness. This result likely reflects the added responsibilities and financial burdens inherent in raising children. Similarly, the coefficient for **Age** (Mean = -0.02) suggests a gradual decline in happiness as individuals age, potentially indicating shifting priorities or life circumstances.

Positive coefficients for **Education Level: Graduate** (Mean = 0.50) and **Job Satisfaction: Moderately Satisfied** (Mean = 0.70) highlight the strong relationship between higher educational attainment, workplace satisfaction, and enhanced happiness. Furthermore, the coefficient for **Real Income** (Mean = 0.02) demonstrates a small but positive effect, affirming the correlation between financial stability and happiness, albeit with diminishing returns.

The Intercept (Mean = -1.50) serves as the baseline log-odds for happiness, providing a reference point against which the impact of other predictors is measured. Collectively, these coefficients clarify the complex interplay of socio-economic and demographic factors in shaping happiness, emphasizing the important role of both external stability and personal well-being in determining life satisfaction.

@@

# Plot the 90% credible intervals for all model coefficients
# This visualizes the impact of each predictor on the ordered priority levels
mcmc_intervals(posterior, prob = 0.9) +
  labs(
    title = "Happiness Model Coefficients",
    subtitle = "90% credible intervals for the predictors",
    x = "Coefficient Estimate",
    y = "Parameters"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5), # Center the main title
    plot.subtitle = element_text(hjust = 0.5) # Center the subtitle
  )
```
As illustrated in @fig-happiness_model_coefficients, the Bayesian logistic regression model provides a clear visualization of the influence of predictor variables on happiness. Each point represents the posterior mean estimate, with horizontal lines denoting the 90% credible intervals, offering a reliable measure of uncertainty. Key findings include:

- The coefficient for **`Marital Status: Married`** is notably positive, indicating that married individuals are significantly more likely to report higher happiness levels compared to those in the reference category (e.g., never married). This finding highlights the importance of marital relationships in fostering well-being.

- **`Education Level: Graduate`** and **`Job Satisfaction: Moderately Satisfied`** exhibit substantial positive effects, underscoring the significant influence of advanced education and workplace satisfaction in enhancing happiness.

- In contrast, the coefficients for **`Number of Children`** and **`Age`** are negative, suggesting that having more children and aging are associated with slight reductions in happiness. These results may reflect the increased financial and caregiving responsibilities tied to larger families, as well as potential shifts in life priorities with age.

- The variable **`Real Income`** demonstrates a modest positive effect, signifying that higher income contributes to increased happiness, albeit with a relatively small magnitude.

Taken together, these findings shed light on the complex interplay of socio-economic and demographic factors shaping individual happiness, providing strong support for targeted interventions aimed at enhancing well-being.

@@

This study investigates the determinants of individual happiness by employing a Bayesian logistic regression model to analyze self-reported happiness levels. Through the careful selection of socio-economic and demographic variables, this research identifies key factors influencing well-being and their relative importance, providing a detailed understanding of the variables shaping happiness.

@@

The analysis prioritizes variables—marital status, job satisfaction, education level, income, age, and number of children—based on their theoretical relevance and empirical association with happiness. This selection ensures that the model reflects real-world dynamics affecting well-being. Findings underscore the strong predictive power of marital status and job satisfaction, as married individuals and those who report high job satisfaction are significantly more likely to describe themselves as "very happy." These results align with established theories emphasizing the role of emotional and economic stability in fostering happiness.

A complex relationship emerges for the number of children, where larger family sizes are associated with marginally lower happiness levels, likely due to increased caregiving and financial responsibilities. Similarly, age shows a slight negative association with happiness, reflecting shifts in life priorities over time. In contrast, higher education levels and income exhibit positive effects, reinforcing the importance of socio-economic stability and opportunities in improving life satisfaction. These findings provide a comprehensive perspective on the relationships among socio-economic and demographic factors influencing happiness.

@@

The deliberate focus on variables with strong theoretical and empirical foundations underscores the diverse factors contributing to happiness. Marital status and job satisfaction emerge as robust predictors, reflecting their significant emotional and economic impact. The inclusion of education and income highlights the role of socio-economic stability in fostering well-being, while the effects of age and family size reveal deeper complexities in how life circumstances shape happiness.

Although this study excludes variables such as occupational categories or regional differences to streamline the model, these factors represent promising avenues for future research. For example, exploring the influence of specific professions or geographic contexts could reveal cultural and environmental interactions with socio-economic predictors. Such investigations would refine current models and expand the understanding of happiness determinants.

@@

Despite its contributions, this study is limited by the scope of variables included and the dataset's size. Excluding variables such as region or health conditions narrows the analysis and leaves room for future research to provide a more holistic perspective. Incorporating additional dimensions, such as leisure activities or access to healthcare, could further enhance the model's explanatory power.

Expanding the dataset to include broader samples across diverse regions or employing longitudinal data could improve generalizability and offer insights into how happiness evolves over time. Moreover, investigating the impact of external socio-economic factors, such as public policy changes or economic crises, could reveal macro-level influences on well-being.

Advancements in analytical methods, such as machine learning techniques, could uncover complex, non-linear relationships among variables that traditional regression models may overlook. For instance, examining interactions between age, income, and job satisfaction could illuminate critical periods and conditions for optimal well-being.

These directions not only address the limitations of this study but also pave the way for a more comprehensive understanding of happiness in varied contexts, advancing both academic and policy-oriented research.

@@

The dataset’s rich potential invites exploration of broader societal influences on happiness. For instance, integrating external datasets to examine the role of social support systems, healthcare access, or environmental quality could illuminate contextual factors that amplify or mitigate individual-level predictors.

Longitudinal analyses offer another promising direction, enabling researchers to track changes in happiness over time and identify key life stages for intervention. Similarly, incorporating geographical data could uncover regional variations in happiness, allowing for nuanced analyses of cultural and environmental determinants.

By building on this study’s findings, future research can deepen our understanding of happiness drivers, enhancing both theoretical frameworks and practical applications aimed at improving well-being.

@@

This study demonstrates the importance of understanding the determinants of happiness. Findings such as the significant roles of marital status and job satisfaction offer actionable recommendations for policymakers and employers to design strategies aimed at enhancing well-being. The positive relationship between higher education and happiness emphasizes the societal importance of investing in equitable education systems, reinforcing the role of socio-economic stability in fostering life satisfaction.

By advancing the understanding of happiness drivers, this research contributes to the broader discourse on well-being and offers a foundation for future studies and interventions aimed at improving individual and societal outcomes.
